---
title: "Arrange Illinois IDEA APR"
author: "Joy Dada"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmdformats::robobook
---

# Introduction

This script arranges the Illinois IDEA Annual Performance Reviews.

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

path_box <- 
  file.path(
    Sys.getenv("BOX"),
    "Child Care and Poverty"
  )

require("pacman")

packages <-
  c("tidyverse",
    "ggplot2",
    "readxl",
    "haven",
    "tidyr",
    "openxlsx",
    "remotes",
    "janitor",
    "expss",
    "dplyr",
    "statar",
    "dataReporter",
    "here",
    "ggeasy",
    "extrafont",
    "sjlabelled",
    "wesanderson",
    "kableExtra",
    "epiDisplay",
    "lfe",
    "stargazer"
  )

pacman::p_load(
  packages,
  character.only = TRUE
)
```


# Import data

## 2014-2017

```{r}
raw_list <-
  lapply(
    1:9,
    function(i){
      lapply(
        2014:2017,
        function(j){
          readRDS(
           here(
            path_box,
            "Data",
            "raw-deidentified",
            "Illinois IDEA",
            j,
            paste0("APR ", j, " ", i, ".rds")
            )
           ) %>% 
            mutate(
              year = j
            ) %>% 
            rename(
              cfc = 1
            )
        }
      )
       
    }
  )
```


# Clean data

Each data frame is different, so we clean them differently

## Delays 2014-2017

### Rename columns

We rename the columns and we label them.

```{r}
delays <-
  reduce(
    raw_list[[1]],
    bind_rows
  ) %>% 
  setnames(
    old = 1:6,
    new = 
      c(
        "cfc",
        "active_ifsp",
        "fec",
        "no_fec",
        "no_delays",
        "pct_no_delays"
      )
  ) 
```


### Fix columns in 2014

In 2014, the % no delays column is split into 2, so we move the values over to the right column and delete the column we don't need.

```{r}
delays <-
  delays %>% 
  mutate(
    pct_no_delays =
      case_when(
        year == 2014 ~ as.character(`...7`),
        TRUE ~ pct_no_delays
      )
  ) %>% 
  dplyr::select(
    - 7
  )
```

### Format columns

For `cfc`, the centers are numbered, but have asterisk next to them to represent which CFC's constitute certain areas:

- 2, 4, 5, 15 & 25 ~ Collar Counties
- 6, 7 & 12 ~ Chicago - Cook County
- 8-11 ~ Suburban - Cook County

We remove the asterisk.

```{r}
delays <- 
  delays %>% 
  mutate(
    cfc =
      trimws(gsub("\\*", "", cfc))
  )
```

Additionally,  we code Statewide, Chicago Cook County, Suburban Cook County and Collar Counties as numbers so we can convert `cfc` to a numbered categorical variable.

```{r, warning = FALSE}
delays <- 
  delays %>% 
  mutate(
    cfc = 
      case_when(
        grepl("Statewide", cfc) ~ "26",
        grepl("Chicago", cfc) ~ "27",
        grepl("Suburban", cfc) ~ "28",
        grepl("\\&", cfc) ~ "29",  # Collar
        grepl("Downstate", cfc) ~ "30",
       TRUE ~ cfc
      ),
    cfc = as.numeric(cfc),
    cfc =
      factor(
        cfc,
        levels = 1:30,
        labels = 
          c(
            as.character(1:25),
            "Statewide",
            "Chicago - Cook County",
            "Suburban - Cook County",
            "Collar Counties",
            "Downstate (All Others)"
          )
      )
  ) 
```

Now we convert all the other columns to numeric.

```{r, warning = FALSE, message = FALSE}
delays <- 
  delays %>% 
  mutate(
    across(
      2:6,
      as.numeric
    )
  )
```

Now we remove rows with missing values.

```{r}
delays <- 
  delays %>% 
  dplyr::filter(
    !(
      is.na(active_ifsp) &
      is.na(fec) &
      is.na(no_fec) &
      is.na(pct_no_delays)
    )
  )
```

### Label variables

```{r}
delays <- 
  delays %>% 
  apply_labels(
    cfc = "Child and Family Connections",
    active_ifsp = "Number of Active Individualized Family Service Plans (IFSP)",
    fec = "Number of Children with Family Exceptional Circumstances",
    no_fec = "Number of Children without Family Exceptional Circumstances",
    no_delays = "Number of Children Recieving EI services on their IFSPs in a timely manner",
    pct_no_delays = "Percentage of Children Recieving EI services on their IFSPs in a timely manner /n (No Delays + FEC)/Active IFSP"
  )
```


### Export data

```{r}
saveRDS(
  delays,
  here(
    path_box,
    "Data",
    "constructed",
    "APR 2014-2017 delays.rds"
  )
)
```


