---
title: "Create CFC County Data"
author: "Joy Dada"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmdformats::robobook
---

# Introduction

This script creates CFC County Data.

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

path_box <- 
  file.path(
    Sys.getenv("BOX"),
    "Child Care and Poverty"
  )

require("pacman")

packages <-
  c("tidyverse",
    "ggplot2",
    "readxl",
    "haven",
    "tidyr",
    "openxlsx",
    "remotes",
    "janitor",
    "expss",
    "dplyr",
    "statar",
    "dataReporter",
    "here",
    "ggeasy",
    "extrafont",
    "sjlabelled",
    "wesanderson",
    "kableExtra",
    "epiDisplay",
    "lfe",
    "stargazer"
  )

pacman::p_load(
  packages,
  character.only = TRUE
)
```

# Import County List

We are using counties listed in the SAIPE.

```{r}
poverty <-
   readRDS(
     here(
       path_box,
       "Data",
       "constructed",
       "SAIPE Poverty County Estimates 2004-2021"
       )
   ) 
```

We look at all counties in Illinois.

```{r}
counties <-
  poverty %>% 
  filter(
    postal == "IL" & county_fips != 0
  ) %>% 
  distinct(name) 
```

There are `r nrow(counties)` counties in Illinois.

# Create data


## Create list of what county each CFC serves

Using [this map](https://providerconnections.org/wp-content/uploads/2021/11/CFC-Map.jpg) and [this table][(https://providerconnections.org/wp-content/uploads/2023/06/2023-Child-Family-Connections-Listing-of-Illinois.pdf).

We first make the county names uniform and then match each county to its CFC.

```{r}
cfc <-
  counties %>% 
  mutate(
    name = trimws(tolower(gsub("County", "", name))),
    cfc =
      case_when(
        name %in% 
          c("boone",
            "bureau",
            "marshall",
            "putnam", 
            "ogle",
            "winnebago")  ~ 1,
        name == "lake" ~ 2,
        name %in%
          c("carroll",
            "dekalb",
            "jo daviess",
            "lee",
            "stephenson",
            "whiteside") ~ 3,
        name %in%
          c("kane", "kendall") ~ 4,
        name == "dupage" ~ 5,
        name == "cook" ~ 6, # 6-12
        name %in% 
          c("fulton",
            "hancock",
            "henderson",
            "knox",
            "mcdonough",
            "mercer",
            "rock island",
            "schuyler",
            "warren") ~ 13,
        name %in% 
          c("henry",
            "peoria",
            "stark",
            "tazewell",
            "woodford") ~ 14,
        name %in% 
          c("grundy",
            "kankakee",
            "la salle",
            "will") ~ 15,
        name %in% 
          c("champaign",
            "ford",
            "iroquois",
            "livingston",
            "mclean",
            "vermilion") ~ 16,
        name %in% 
          c("adams",
            "brown",
            "calhoun",
            "cass",
            "greene",
            "jersey",
            "morgan",
            "pike",
            "scott") ~ 17,
        name %in% 
          c("logan",
            "menard",
            "mason",
            "sangamon") ~ 18,
        name %in%
          c("clark",
            "coles",
            "cumberland",
            "de witt",
            "douglas",
            "edgar",
            "macon",
            "moultrie",
            "piatt",
            "shelby") ~ 19,
        name %in%
          c("bond",
            "christian",
            "clay",
            "crawford",
            "effingham",
            "fayette",
            "jasper",
            "lawrence",
            "macoupin",
            "montgomery",
            "richland") ~ 20,
        name %in%
          c("madison",
            "monroe",
            "randolph",
            "st. clair") ~ 21, # flag
        name %in%
          c("clinton",
            "marion",
            "franklin",
            "jefferson",
            "washington",
            "williamson") ~ 22,
        name %in% 
          c("edwards",
            "gallatin",
            "hamilton",
            "saline",
            "wabash",
            "wayne",
            "white") ~ 23,
        name %in% 
          c("alexander",
            "hardin",
            "jackson",
            "massac",
            "perry",
            "pope",
            "pulaski",
            "johnson",
            "union",
            "massac") ~ 24,
        name == "mchenry" ~ 25,
        TRUE ~ NA_real_
        )
      )
```

## Add other CFCs for Cook County

Since Cook County is served by CFCs 6-12, we must add entries for CFC 7-12.

```{r}
cfc <-
  cfc %>% 
  add_row(
    name = rep("cook", 6),
    cfc = 7:12
  )
```

### Tabulate CFCs

We do this to check every county has been assigned a CFC.

```{r}
tab1(cfc$cfc)
```

# Export data

```{r}
saveRDS(
  cfc,
  here(
    path_box,
    "Data",
    "constructed",
    "CFC_county_master.rds"
   )
)
```
