---
title: "Missouri Caseload Counter"
author: "Joy Dada"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmdformats::robobook
---

# Introduction

This script imports and analyzes the [Missouri Department of Social Services Caseload Counter](https://dss.mo.gov/mis/clcounter/history.htm).

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Change to Box folder on your computer

path_box <- 
  file.path(
    Sys.getenv("BOX"),
    "Child Care and Poverty"
  )

# Load necessary packages

require("pacman")

packages <-
  c("tidyverse",
    "ggplot2",
    "bife",
    "readxl",
    "haven",
    "tidyr",
    "openxlsx",
    "remotes",
    "janitor",
    "expss",
    "dplyr",
    "statar",
    "dataReporter",
    "here",
    "ggeasy",
    "extrafont",
    "sjlabelled",
    "wesanderson",
    "kableExtra",
    "epiDisplay",
    "lfe",
    "stargazer"
  )

pacman::p_load(
  packages,
  character.only = TRUE
)
```


# Import data

```{r, warning = FALSE}
caseload_clean <-
  readRDS(
    here(
      path_box,
      "Data",
      "constructed",
      "MO Caseload Counter 2004-05 to 2023-08"
    )
  )
```

## Glimpse of data

```{r}
head(caseload_clean, 5) %>% 
  kable(
    .,
    format = "html",
    caption = "MO Caseload Counter 2004-05 to 2023-08",
    align = "c"
  ) %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```

# Import historical population data

```{r}
hist_clean <-
  readRDS(
    here(
          path_box,
          "Data",
          "constructed",
          "missouri-population-2004-2022.rds"
        )
  )
```

## Glimpse of data

```{r}
head(hist_clean, 5) %>% 
  kable(
    .,
    format = "html",
    caption = "Glimpse of clean historical population data",
    align = "c"
  ) %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```

# Import historical child population data

```{r}
child_clean <-
  readRDS(
    here(
      path_box,
      "Data",
      "constructed",
      "Child population by age group.rds"
    )
  )
```


# Children Receiving Subsidized Child Care 

```{r}
childcare <-
  caseload_clean %>% 
  filter(
    variable %in%
      c("Children Receiving Subsidized Child Care",
        "Pregnant Women")
  )
```

## Add dummy variable for the introduction of the abortion ban

```{r}
childcare <-
  childcare %>% 
  mutate(
    ban = 
      case_when(
        year >= 2022 & month > 6 ~ 1,
        TRUE ~ 0
      ),
    ban = 
      factor(
        ban,
        levels = 0:1,
        labels = c("No ban", "Ban")
      )
  ) %>% 
  pivot_wider(
    names_from = variable,
    values_from = count
  ) %>% 
  setnames(
    old = 5:6,
    new = c("preg", "childcare_sub")
  ) %>% 
  apply_labels(
    preg = "Pregnant Women",
    childcare_sub = "Children Receiving Subsidized Child Care"
  ) %>%
  group_by(year) %>% 
  mutate(
    date_fe = cur_group_id()
  )  %>% 
  ungroup()

tab1(childcare$ban)
```

## Plot trend in 2022 for the number of active cases

```{r}
childcare %>%
  filter(year >= 2019) %>% 
  ggplot() +
  geom_point(
    aes(
      x = date,
      y = childcare_sub
    ) 
  ) + 
  theme_bw() +
  geom_vline(
    xintercept = "2022-06"
  ) 
  # scale_x_continuous(
  #   breaks = 1:12
  # ) 
```


## Quick regression

No fixed effects

```{r}
model1 <-
  lm(
    childcare_sub ~ ban  + preg,
    data = childcare
  )

stargazer(
  model1,
  type = "text"
  )
```

### Yearly fixed effects

The ban decreases the the number of children recieving subsidized child care.

```{r}
model2 <-
  felm(
    childcare_sub ~ ban + preg | date_fe,
    data = childcare
  )

stargazer(
  model2,
  type = "text"
  )
```


## Merge population with counts

```{r}
childcare_merge <-
  childcare %>% 
 left_join(
    hist_clean,
    by = "year"
  ) %>% 
  arrange(year) %>% 
  fill(tot_pop)
```

## Compute share of population that are pregnant women

```{r}
childcare_pop <-
  childcare_merge  %>% 
  mutate(
    pct_preg = preg/tot_pop
  )
```

# Merge population with child care count

```{r}
childcare_df <-
  childcare_pop %>% 
  left_join(
    child_clean,
    by = "year"
  )  %>% 
  arrange(year) %>% 
  fill(
    child_pop,
    .direction = "down"
    ) 
```

## Create percentage of children recieiving child care subsidies

```{r}
childcare_df <-
  childcare_df %>% 
  mutate(
    pct_child = childcare_sub/child_pop
  ) %>% 
  group_by(
    year
  ) %>% 
  mutate(
    date_fe = cur_group_id()
  ) %>% 
  ungroup() %>% 
  mutate(
    date = as.Date(paste0(date, "-01"))
  )
```

## Scatter plot with `pct_child`

Shows a clear downward trend since 2004

```{r}
childcare_df %>% 
  ungroup() %>% 
  ggplot(
    aes(
      x = date,
      y = pct_child
    )
  ) +
  geom_point(
    size = 0.5
  ) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::percent
    ) +
  scale_x_date(
    date_breaks = "2 year",
    date_labels = "%Y"
  ) +
  geom_smooth( 
    method = 'lm'
    ) +
  labs(
    title = "% of Children Recieving \n Subsidized Child Care",
    y = "",
    x = "Date",
    subtitle = "Missouri"
  ) +
  theme(
    plot.title = 
      element_text(
        hjust = 0.5,
        face = "bold",
        size = 12),
    plot.subtitle =
      element_text(
        hjust = 0.5,
        face = "italic",
        size = 11
      )
  )
```

## Scatter with `pct_preg`

```{r}
childcare_df %>% 
  ungroup() %>% 
  ggplot() +
  geom_point(
    aes(
      x = date,
      y = pct_preg
    )
  ) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::percent
    )
```

## Regression with time fixed effects

```{r}
model3 <-
  felm(
    pct_child ~ ban + pct_preg | date_fe,
    data = childcare_df
  )

stargazer(
  model3,
  type = "text"
)
```

Negative effect on the share of children recieving subsidised child care, controlling for the percentage of pregnant women in the population.

