---
title: "Geocode Illinois CFCs"
author: "Joy Dada"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmdformats::robobook
---

# Introduction

This script geocodes the locations of Illinois 25 CFCs.

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

path_box <- 
  file.path(
    Sys.getenv("BOX"),
    "Child Care and Poverty"
  )

require("pacman")

packages <-
  c("tidyverse",
    "leaflet",
    "janitor",
    "readxl",
    "tidyr",
    "kableExtra",
    "tidygeocoder",
    "sf",
    "here",
    "ggmap",
    "map",
    "mapdata",
    "ggplot2",
    "googleway"
    # "openxlsx",
    # "remotes",
    # "expss",
    # "statar",
    # "dataReporter",
    # "here",
    # "ggeasy",
    # "extrafont",
    # "sjlabelled",
    # "wesanderson",
    # "kableExtra",
    # "epiDisplay",
    # "lfe",
    # "stargazer"
  )

pacman::p_load(
  packages,
  character.only = TRUE
)
```


# Import data

We will start of using a spreadsheet including all CFCs, and their addresses.

```{r}
raw_cfc <-
  readRDS(
    here(
        path_box,
        "Data",
        "constructed",
        "CFC Address Master.rds"
      )
  )
```

# Geocode addresses

We use the package `tidygeocoder` to geocode the addresses. We use [OpenStreetMap data](https://nominatim.org/) to geocode, through choosing the `osm` method. This process will take roughly 30 seconds.

```{r, message = FALSE}
coord <-
  raw_cfc %>% 
  mutate(
    address =
      paste(
        street_address,
        postal_code
      )
  ) %>% 
  tidygeocoder::geocode(
    address = street_address,
    method = 'osm',
    lat = latitude ,
    long = longitude
  )


api_key <-
  "AIzaSyDLMb3IZXkLXzvjVDfMhrAw4usUfHQvShc"

geocode_search <- function(list){
  
  lapply(seq_along(list), function(i){
    
    raw_output <- 
      google_geocode(
        address = list[[i]],
        key = api_key,
        simplify = TRUE)$results$geometry$location
    
    if(is.null(class(raw_output))){
      
      print(paste0("No coordinates found for ",  list[[i]]))
      
      results <- data.frame(
        lat = NA_real_,
        lng = NA_real_,
        search_name = list[[i]]
      ) 
      
      results
      
    }
    
    else {
      
      print(paste0("Coordinates found for ",  list[[i]]))
      
      raw_output$search_name =  list[[i]]
      
      
      raw_output
    
    }
    
  })
  
}

geocode_raw_results <- 
  geocode_search(unique(coord$street_address))

```

For addresses that that were not able to find an address for, we search using their postal code. This will take roughly 20 seconds.

```{r, message = FALSE}
missing_coord <-
  coord %>% 
  filter(
    is.na(latitude)
  ) %>% 
  select(-starts_with("l")) %>% 
  mutate(
    postal_code = substr(postal_code, 1, 8),
    postal_code =
      gsub("IL", "Illinois", postal_code),
    address = 
      paste(
        street_address, postal_code
      )
  )


```




```{r}


  tidygeocoder::geocode(
    address = postal_code,
    method = 'osm',
    lat = latitude ,
    long = longitude
    )
```

Resolve issue of duplicate coordinates i.e. multiple CFCs having the same coordinate.

We create an indicator column `type` to tell us if we found the coordinate through the street address or through its postcode. We combine all coordinates into 1 data frame.

```{r}
all_coord <-
  missing_coord %>% 
  mutate(
    type = "postal"
  ) %>% 
  bind_rows(
    filter(
      coord,
      !is.na(latitude)
    )
  ) %>% 
  mutate(
    type = 
      ifelse(
        is.na(type),
        "street",
        type
      ),
    across(
      starts_with("l"),
      as.numeric
    )
  )


all_coord %>% 
  tabyl(type)
```

## Convert to sf object

We convert the coordinates to coordinates in a sf object.

```{r}
gps <-
  all_coord %>% 
  st_as_sf(
    coords = c("longitude", "latitude"),
    crs = 4326
  )
```


# Export data

```{r}
saveRDS(
  gps,
  here(
    path_box,
    "Data",
    "tidy",
    "IL_CFC_coord.rds"
  )
)
```

# Map of CFCs

```{r}
map_df <-
  gps %>% 
  mutate(
    info = paste(
      "Location name", name, "<br>",
      "CFC #", cfc, "<br>"
    ),
    cfc = factor(cfc)
  )

cols <-
  rainbow(
    length(
      levels(
        map_df$cfc
      )
    ),
    alpha = NULL
  )
  
map_df$colors <-
  cols[unclass(map_df$cfc)]
  
il_base <-
  map_data("state") %>% 
  filter(
    region == "illinois"
  )

il_county <-
  map_data("county") %>%  
  filter(
    region == "illinois"
  )

  ggplot(
    data = all_coord,
    aes(
      x = longitude,
      y = latitude,
      # color = factor(cfc),
      label = cfc
      ), 
     inherit.aes = FALSE,
  ) +
  geom_point(
    size = 10,
    fill = "white",
    color = "white"
  ) +
  geom_text(
    vjust = 0.4,
    color = "black"
  ) +
  coord_fixed(1.3) +
  guides(
    color = "none",
    size = "none"
    ) +
  coord_fixed(1.3) + 
  geom_polygon(
  data = il_base,
  aes(x = long, y = lat, group = group),
  color = "black", fill = NA) + 
  theme_bw() +
  coord_fixed(1.3) + 
  geom_polygon(
    data = il_county,
    fill = NA, 
    color = "black"
    ) +
  geom_polygon(
    color = "black",
    fill = NA
    ) 
  
  
  
  scale_color_manual(
    tile = "Child and Family Connection Centers"
  )
  
  


leaflet() %>%
  addProviderTiles(
    'Esri.WorldImagery',
    options = providerTileOptions(opacity = .9)
  ) %>%
  addCircleMarkers(
    data = map_df,
    label = lapply(map_df$info, htmltools::HTML),
    radius = 5,
    color = ~ colors,
    opacity = .9
  ) %>% 
  setView(
    lng = -89,
    lat = 40,
    zoom = 5
    ) 
```

