---
title: "Import and arrange historical poverty data"
author: "Joy Dada"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmdformats::robobook
---

# Introduction

This script imports and arrange historical poverty data from the [US Census Bureau](https://www.census.gov/library/publications/2023/demo/p60-280.html#:~:text=The%20SPM%20rate%20in%202022,and%20Table%20B%2D3).

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Change to Box folder on your computer

path_box <- 
  file.path(
    Sys.getenv("BOX"),
    "Child Care and Poverty"
  )

# Load necessary packages

require("pacman")

packages <-
  c("tidyverse",
    "ggplot2",
    "readxl",
    "haven",
    "tidyr",
    "openxlsx",
    "remotes",
    "janitor",
    "expss",
    "dplyr",
    "statar",
    "dataReporter",
    "here",
    "ggeasy",
    "extrafont",
    "sjlabelled",
    "wesanderson",
    "kableExtra",
    "epiDisplay",
    "lfe",
    "stargazer"
  )

pacman::p_load(
  packages,
  character.only = TRUE
)
```


# Import poverty data

```{r, warning = FALSE, message = FALSE}
poverty_list <-
    lapply(
    4:21,
    function(year){
      
      read_excel(
            here(
              path_box,
              "Data",
              "raw-deidentified",
              "State and County Poverty Estimates",
              paste0("est", year, "all.xls")
          ),
          skip = 1
        ) %>% 
        mutate(
          year = year
        )
    }
    )
```

## Glimpse at data


```{r}
head(poverty_list[[1]], 5) %>% 
  kable(
    .,
    format = "html",
    caption = "Raw US Census Bureau Poverty Data",
    align = "c"
  ) %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```


# Clean data

Each year has data organised slightly differently so we clean them differently.

## 2004

```{r}
poverty_04 <-
  poverty_list[[1]] %>%
  clean_names() 
```

### Select variables and rename

```{r}
poverty_04 <-
  poverty_04 %>% 
  dplyr::select(
    name,
    county_fips,
    year,
    postal,
    poverty_percent_all_ages,
    poverty_percent_under_age_18,
    poverty_percent_ages_5_17
  ) %>% 
  rename_with(
    ~ gsub("_percent", "", .x),
    starts_with("poverty_")
  )
```


### Select state level variables

```{r}
poverty_04 <-
  poverty_04 %>%
  filter(
    county_fips == 0
  ) %>% 
  dplyr::select(
    -county_fips
  )
```

### Convert to numeric

```{r}
poverty_04 <-
  poverty_04 %>%
  mutate(
    across(
      starts_with("poverty_"),
      as.numeric
    )
  ) 
```

## 2005-2012

These years have the same structure so we clean these at the same time.

First we bind the rows.

```{r}
poverty_0512 <-
  reduce(
   poverty_list[2:9],
   bind_rows
  )
```

### Convert 1st row to column names 

```{r, warning = FALSE}
poverty_0512 <-
  poverty_0512 %>% 
  row_to_names(
    row_number = 1
  ) %>% 
  clean_names()
```

### Select variables and rename

```{r}
poverty_0512 <-
  poverty_0512 %>% 
  dplyr::select(
    name,
    county_fips,
    x5,
    postal,
    poverty_percent_all_ages,
    poverty_percent_under_age_18,
    poverty_percent_ages_5_17
  ) %>% 
  rename_with(
    ~ gsub("_percent", "", .x),
    starts_with("poverty_")
  ) %>% 
  rename(
    year = x5
  )
```

### Select state level variables

```{r}
poverty_0512 <-
  poverty_0512 %>% 
  filter(
    county_fips == 0
  ) %>% 
  dplyr::select(
    -county_fips
  )
```

### Convert to numeric

```{r}
poverty_0512 <-
  poverty_0512 %>% 
  mutate(
    across(
      starts_with("poverty_"),
      as.numeric
    )
  ) 
```

## 2013-2021

These years have the same structure so we clean these at the same time.

First we bind the rows.

```{r}
poverty_1321 <-
  reduce(
   poverty_list[10:18],
   bind_rows
  )
```


### Convert 2nd row to column names 

```{r, warning = FALSE}
poverty_1321 <-
  poverty_1321 %>% 
  row_to_names(
    row_number = 2
  ) %>% 
  clean_names()
```

### Select variables and rename

In the previous years of data the following variables mirror:
- `poverty_percent_under_age_18` ~ `poverty_percent_age_0_17`
- `poverty_percent_ages_5_17` ~ `poverty_percent_ages_5_17_in_families`

We rename them for consistency and ease of merging.

```{r}
poverty_1321 <-
  poverty_1321 %>%  
  dplyr::select(
    name,
    county_fips_code,
    x13,
    postal_code,
    poverty_percent_all_ages,
    poverty_percent_age_0_17,
    poverty_percent_age_5_17_in_families
  ) %>% 
  rename_with(
    ~ gsub("_percent", "", .x),
    starts_with("poverty_")
  ) %>% 
  setnames(
    old = c(
      "x13",
      "postal_code",
      "poverty_age_0_17",
      "poverty_age_5_17_in_families"
    ),
    new = c(
      "year",
      "postal",
      "poverty_under_age_18",
      "poverty_ages_5_17"
    )
  )
```

### Convert to numeric

```{r, warning = FALSE}
poverty_1321 <-
  poverty_1321 %>%  
  mutate(
    across(
      starts_with(
        c("poverty_",
          "county_"
          )
        ),
      as.numeric
    )
  ) 
```


### Select state level variables

```{r}
poverty_1321 <-
  poverty_1321 %>%  
  filter(
    county_fips_code == 0
  ) %>% 
  dplyr::select(
    -county_fips_code
  )
```

## Merge yearly data

```{r}
poverty_raw <-
  reduce(
    list(
      poverty_04,
      poverty_0512,
      poverty_1321
    ),
    bind_rows
  )
```

## Fix year indicator after merge

```{r}
poverty <-
  poverty_raw %>%
  rename(
    year_ = year
  ) %>% 
  mutate(
    date = 
      case_when(
        str_length(
          as.character(year_)
          ) == 1 ~ paste0("200", year_,"-01-01"),
        str_length(
          as.character(year_)
          ) == 2 ~ paste0("20", year_, "-01-01")
      ),
    date = as.Date(date),
    year = as.numeric(substr(date, 1, 4))
  )  %>% 
  dplyr::select(-year_) 

tab1(poverty$year)
```

## Convert percentages to decimals

```{r}
poverty_clean <-
  poverty %>% 
  mutate(
    across(
      starts_with("poverty_"),
      ~ .x/100
    )
  )
```

## Export 

```{r}
saveRDS(
  poverty_clean,
  here(
    path_box,
    "Data",
    "constructed",
    "SAIPE Poverty Estimates 2004-2021"
  )
)
```

# Descriptive statistics

Focus on Missouri poverty trends overtime.

## Prepare data

Pivot longer so we can plot all variables in the same plot.

```{r}
poverty_df <-
  poverty %>% 
  pivot_longer(
    cols = 3:5,
    names_to = "group",
    values_to = "pct_poverty"
  )  %>% 
  mutate(
    group = 
      factor(
        group,
        levels = 
          c(
            "poverty_all_ages",
            "poverty_ages_5_17",
            "poverty_under_age_18"
            ),
        labels =
          c(
            "All ages",
            "Ages 5-7",
            "Under 18"
          )
      )
  )
```


## Plot over time

```{r, fig.align = 'center', fig.height = 4, fig.width = 7}
cols <- 
  c(
    "All ages" = "pink",
    "Ages 5-7" =  "lightblue", 
    "Under 18" = "purple3"
    )

poverty_df %>% 
  filter(
    postal == "MO"
  ) %>% 
  ggplot(
    aes(
      y = pct_poverty,
      x = year,
      color = group,
      group = group
      )
  ) +
  geom_point(
    size = 2
  ) +
  geom_line(
    linewidth = 1.5
  ) +
  theme_bw() +
  scale_x_date(
    date_breaks = "1 year",
    date_labels = "%Y"
  ) +
  labs(
    x = "Year",
    y = "% living in poverty",
    title = "Percentage of Missourians living in poverty by group",
    subtitle = "2004-2021",
    color = ""
  ) +
  theme(
    plot.title =
      element_text(
        hjust = 0.5,
        size = 12,
        face = "bold"
      ),
    plot.subtitle =
      element_text(
        hjust = 0.5,
        size = 11,
        face = "italic"
      ),
    legend.position =
      c(.9,.87),
    legend.background = element_blank(),
    legend.key = element_blank()
  ) +
  scale_color_manual(
    values = cols
  ) +
  scale_y_continuous(
    limits = c(12, 23),
    breaks = seq(12, 23, 2)
  ) 

ggsave(
  filename =
    here(
      path_box,
      "Output",
      "Poverty in Missouri 2004-21.jpeg"
    ),
  height = 4,
  width = 7
)
```