---
title: "Merge Illinois APR and SAIPE"
author: "Joy Dada"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmdformats::robobook
---

# Introduction

This script merges the Illinois APR CFCs and the SAPIE Poverty Data.

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

path_box <- 
  file.path(
    Sys.getenv("BOX"),
    "Child Care and Poverty"
  )

require("pacman")

packages <-
  c("tidyverse",
    "ggplot2",
    "readxl",
    "haven",
    "tidyr",
    "openxlsx",
    "remotes",
    "janitor",
    "USpopcenters",
    "tigris",
    "expss",
    "dplyr",
    "statar",
    "leaflet",
    "dataReporter",
    "here",
    "ggeasy",
    "extrafont",
    "sjlabelled",
    "wesanderson",
    "kableExtra",
    "epiDisplay",
    "lfe",
    "stargazer",
    "sf"
  )

pacman::p_load(
  packages,
  character.only = TRUE
)
```


# Import data

## 2014-2017 Illinois APR

```{r}
delays <-
  readRDS(
    here(
    path_box,
    "Data",
    "constructed",
    "APR 2014-2017 delays.rds"
    )
  )
```

## CFC County Master

```{r}
cfc <-
  readRDS(
    here(
    path_box,
    "Data",
    "constructed",
    "CFC_county_master.rds"
   )
)
```

## SAIPE

```{r}
poverty <-
   readRDS(
     here(
       path_box,
       "Data",
       "constructed",
       "SAIPE Poverty County Estimates 2004-2021"
       )
   )
```

# Arrange data

## County Spatial Data

We import polygons for Illinois counties using the `tigris` package, which fetches the data from 2010, and we format county name.

```{r, message = FALSE}
spatial <-
  counties(state = "Illinois") 

spatial <- 
 spatial %>% 
  clean_names() %>% 
  dplyr::select(
   name, statefp, countyfp 
  ) %>% 
  mutate(
     name = trimws(tolower(gsub("County", "", name))),
  )
```



## SAIPE

We will restrict the data to Illinois and format the county names to match the `cfc_county_master` data.

```{r}
pov_il <-
  poverty %>% 
  filter(
    postal == "IL"
  ) %>% 
  mutate(
   name = trimws(tolower(gsub("County", "", name))),
  )
```

## Illinois APR delays

We will convert the `cfc` column to numeric to prepare for the later merge. We also remove the variable labels as they interrupt the merge.

```{r, warning = FALSE}
delays_clean <-
  delays %>% 
  mutate(
    cfc = as.numeric(cfc)
  ) %>% 
  zap_labels()
```

# Merge data
## Merge CFC and SAIPE Estimates

We code Illinois as CFC 26 so we can merge the statewide variables from the APR delay tables.

```{r}
pov_cfc <-
  cfc %>% 
  add_row(
    name = "illinois",
    cfc = 26
  ) %>% 
  inner_join(
    pov_il,
    by = "name",
    relationship = "many-to-many"
  )
```

## Merge Illinois APR Tables

```{r}
pov_cfc_apr <-
  pov_cfc %>% 
  inner_join(
    delays_clean,
    by = c("cfc", "year")
  )
```

## Merge spatial county data

```{r}
spatial_mrg <-
  pov_cfc_apr %>% 
  inner_join(
    spatial,
    by = "name"
  )
```

## Make data set with the CFC's coordinates


# Export data

```{r}
saveRDS(
  spatial_mrg,
  here(
    path_box,
    "Data",
    "constructed",
    "Delays and Poverty (2014-2017).rds"
  )
)
```


# Descriptive statistics

```{r}
map_dta <-
  spatial_mrg %>% 
    filter(
      year == 2014
     ) %>% 
    mutate(
      pov_quant = xtile(poverty_under_age_18, 5),
      pov_quant =
        factor(
          pov_quant,
          levels = 1:5,
          labels = c("Q1: Relatively rich", "Q2", "Q3", "Q4", "Q5: Relatively poor")
        ),
      info = paste(
        "County name:", name %>% as.character(), "<br>",
        "CFC:", cfc %>% as.character(), "<br>",
        "Child Poverty Rate:", pov_quant %>% as.character(), "<br>"
      ),
      name = factor(name)
    ) %>% 
  st_as_sf() %>% 
  st_transform(crs = 4326)

cols <-
  rainbow(
    length(
      levels(
        map_dta$pov_quant
      )
    ),
    alpha = NULL
  )

map_dta$colors <-
  cols[unclass(map_dta$pov_quant)]

leaflet() %>%
  addProviderTiles(
    'Esri.WorldImagery',
    options = providerTileOptions(opacity = .9)
  ) %>%
  addPolygons(
    data = map_dta,
    label = lapply(map_dta$info, htmltools::HTML),
    color = ~ colors,
    opacity = .9
  )  %>% 
  addLegend('bottomright',
            colors = cols,
            labels =
              c("Q1: Relatively rich",
                "Q2",
                "Q3", 
                "Q4", 
                "Q5: Relatively poor"),
            title = 'Child Poverty Rate (2014)',
            opacity = 1)
```
